Reasoning Graphically about Quantum Computation

### 2008-02-26 papers/reasoning-graphically/paper.tex
working on !box graphs
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1374 f8f375fb-4f10-0410-8d47-b6fbdc7ca053
adding th start of the graphs stuff
git-svn-id: https://isaplanner.svn.sourceforge.net/svnroot/isaplanner/trunk/quantomatic@1375 f8f375fb-4f10-0410-8d47-b6fbdc7ca053

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[runningheads]{llncs}
%\usepackage[dvips]{color}
%\usepackage{multicol}
%\usepackage{tabls}
%\usepackage{ttbox}
\usepackage{allmtt}
%\usepackage{pstricks,pst-node,pst-tree} % PS Tricks for diagrams
%\usepackage{proof2}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{stmaryrd}
%\usepackage{mbboard}
\usepackage{xspace}
%\usepackage{diagrams}

@@

%%--------------------------------------------

@@

@@

\begin{definition}
  A \emph{directed graph} consists of a 4-tuple $(N,E,s,t)$ where $V$
  and $E$ are sets, respectively of \emph{nodes} and \emph{edges}, 
\end{definition}

@@

\subsubsection{}

>>>>

%%%%%%%%%%%%%%%%%%%%%%% file typeinst.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is the LaTeX source for the instructions to authors using
% the LaTeX document class 'llncs.cls' for contributions to
% the Lecture Notes in Computer Sciences series.
% http://www.springer.com/lncs       Springer Heidelberg 2006/05/04
%
% It may be used as a template for your own input - copy it
% to a new file with a new name and use it as the basis
% for your article.
%
% NB: the document class 'llncs' has its own and detailed documentation, see
% ftp://ftp.springer.de/data/pubftp/pub/tex/latex/llncs/latex2e/llncsdoc.pdf
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\documentclass[runningheads]{llncs}
%\usepackage[dvips]{color}
%\usepackage{multicol}
%\usepackage{tabls}
%\usepackage{ttbox}
\usepackage{allmtt}
%\usepackage{pstricks,pst-node,pst-tree} % PS Tricks for diagrams
%\usepackage{proof2}
\setcounter{tocdepth}{3}
\usepackage{graphicx}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{latexsym}
\usepackage{stmaryrd}
%\usepackage{mbboard}
\usepackage{xspace}
\usepackage{diagrams}

@@

%%%% More handy things for me - rwd
\newcommand{\TODO}[1]{%
\typeout{WARNING!!! there is still a TODO left}
\marginpar{\textbf{!TODO: }\emph{#1}}
}
%%--------------------------------------------

@@

Want to do reasoning about quantum computation by manipulating graphs

CAn formalise quantum mechanics in an abstract setting using compact
or $\dag$-compact categories.

We will give a description of graphs in general;  a specialisation of
this gives the structure required for compact closed  categories.  The
graph structure suffices to give a tight representation.

We then introduce a concrete set of generators and equations for use
in quantum computation.  We need rewrites to capture the non-logical
axioms of this structure.   It turns out that we can most easily express
these equations using an infinite family of ``spider rules'';  this
motivates the  more general notion of pattern graphs.

next section introduces the  notion of pattern graph, which is
essentially a graph structured

@@

\subsection{Graphs}
\label{sec:graphs}

A \emph{directed graph}\footnote{
Equivalently:  a directed graph is a functor $G$ from  $\bullet
\pile{\rTo\\\rTo} \bullet$ to \catSet;  a graph morphism is then a natural
transformation  $f: G \Rightarrow H$.
} consists of a 4-tuple $(V,E,s,t)$ where $V$
and $E$ are sets, respectively of \emph{vertices}\footnote{We will use
  the words ``vertex'' and ``node'' interchangeably.} and \emph{edges},
and $s$ and $t$ are maps 
\begin{diagram}
  E & \pile{\rTo^{\qquad s\qquad}\\\rTo_t} & V
\end{diagram}
which we call \emph{source} and \emph{target}.  The \emph{in-degree}
of a vertex $v$ is the cardinality of $t^{-1}(v)$; its out-degree is
the cardinality of $s^{-1}(v)$.  The \emph{degree} of a vertex is
the sum of its in- and out-degrees.

Given graphs $G$ and $H$, a \emph{graph morphism} $f : G\to H$ consists of
functions $f_E : E_G \to E_H$ and $f_V:V_G\to V_H$ such that:
\begin{gather*}
  s_H\circ f_E = f_V \circ s_G,\\
  t_H\circ f_E = f_V \circ t_G.
\end{gather*}
\TODO{different kinds of morphism?}

Augmented by some additional structure, graphs form a compact closed
category.  We present an overview of this structure:  the details
omitted here can be found in \cite{Duncan:thesis:2006}.

A \emph{concrete graph} $\Gamma$ is 5-tuple $(G, \dom\Gamma, \cod\Gamma,
<_{\text{in}(\cdot)}, <_{\text{out}(\cdot)})$ where:
\begin{itemize}
\item $G = (V,E,C,s,t)$ is a graph;
\item $\dom\Gamma$ and $\cod\Gamma$ are totally ordered disjoint of
  degree one vertices of $G$.
\item $<_{\text{in}(\cdot)}$ is a family of maps, indexed by $V$ such
  that $<_{\text{in}(v)} : \text{in}(v) \rTo^\isomorphism
  \mathbb{N}_k$ where $k = \sizeof{\text{in}(v)}$.
\item $<_{\text{out}(\cdot)}$ is a family of maps, indexed by $V$ such
  that $<_{\text{out}(v)} : \text{out}(v) \rTo^\isomorphism
  \mathbb{N}_{k'}$ where $k' = \sizeof{\text{out}(v)}$.
\end{itemize}

Since the sets $\dom\Gamma$ and $\cod\Gamma$ consist of vertices of degree
one, we can assign a polarity to each one:  $v \mapsto +$ if the edge
incident at $v$ is an incoming edge; $v \mapsto -$ otherwise.  Hence
$\cod \Gamma$ and $\dom \Gamma$ are \emph{signed sets}.  Concrete
graphs form a category whose objects are

can be seen as giving a signing to each element of the ordinal
one, we can view these sets as maps from the ordinal

can be viewed as a maps from an

@@

\subsubsection{Matching}

\subsubsection{Plugging}
